---
title: "4. C-style Formatting"
date: "2018-09-25"
output:
  rmdformats::html_docco:
    highlight: kate
---


```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)
library(magrittr)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
               cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

`r Sys.Date()` edited

noted by @kemushi54

<br>

# C-style format是啥？
C-style Formating: formatting like in the C language
function name in other language: printf (print formattinf)

# When would you use `sprintf()`?
  * exporting output to some file.
  * printing output on console.
  * forming new strings.

```{r}
sprintf("I woke up at %s:%s%s a.m.", 8, 0, 5)
# %: as a slot, is a placeholder for a variable that will be formatted
# s: for type of the original vector, and how it will be formatted

hour <- 8
mins1 <- 0
mins2 <- 5
sprintf("I woke up at %s:%s%s a.m.", hour, mins1, mins2)
```

<br>

# 4.1 C-style formatting options
* string: %s
* interger: %d, %0xd
* number: %f, %.xf
* scientific notation: %e, %E, %g

<br>

## 4.1.1 Format Slot Syntax
The full syntax for a format slot:

* **%[parameter][flags][width][.precision][length]type**

```{r}
# parameter, n$ (%n$d)
sprintf("%d, %d", 2, 1)
sprintf("%2$d, %2$d, %1$d", 2, 1)
```

<br>
```{r}
# flags (+, -, ' ', 0, #)
# "+" 代表顯示正負號
# "-" 代表靠左對齊
sprintf("%-d, -%d, %+d,  %d", 1, 2, 3, 4)
sprintf(' "%d" ', 2)
sprintf("'%d'", 2)

sprintf("%g", 1.2340000)
sprintf("%#g", 1.2340000)
sprintf("%g", 1234)
sprintf("%#g", 1234)
```

```{r}
# width and precision 
  # width: minimum number of characters to output
    # does not cause truncation of oversized fields
  # precision: maximum limit on the output
  # interger
sprintf("%*d", 5, 10)
sprintf("%5d", 10)
sprintf("%.5d", 10)
sprintf("%05d", 10)
  # number (default of precision = 6)
sprintf("%*f", 5, 10)
sprintf("%5f", 10)
sprintf("%.2f", 10)
sprintf("%05.2f", 10)
sprintf("%08.2f", 10)
sprintf("%8.2f", 10)
```

```{r}
  # character
sprintf("%.*s", 3, "abcdef")
sprintf("%*s", 3, "abcdef")
```

```{r}
# length ??? 課本沒寫到

# type 見課本
```

<br>

## 4.1.2 Example: basic `sprintf()`
```{r}
sprintf("%s", 1/6)
sprintf("%f", 1/6)
sprintf("%.3f", 1/6)
sprintf("%e", 1/6)
sprintf("%.2e", 1/6)
sprintf("%g", 1/6)
```

<br>

## 4.1.3 Example: File Names
```{r}
paste0("data", 1:10, ".csv")
sprintf("data%02d.csv", 1:10)
```

<br>

## 4.1.4 Example: Fahrenheit to Celsius
```{r}
# function to_celsius()
to_celsius <- 
  function(temp = 1) {
    (temp - 32) * 5/9
}

to_celsius(95)

fahrenheit2celsius <- function(temp = 1) {
  celsius <- to_celsius(temp)
  sprintf('%.2f Fahrenheit degrees = %.2f Celsius degrees', temp, celsius)
}

fahrenheit2celsius(95)
```

<br>

## 4.1.5 Example: Car Traveled Distance
* creat S3 class object
* https://www.cyclismo.org/tutorial/R/s3Classes.html
* http://map-testing.blogspot.com/2016/09/r-s3s4.html

```{r}
mazda3 <- list(
  name = 'mazda3', # car name
  make = 'mazda',  # car make
  year = 2015,     # year model
  city = 30,       # fuel consumption in city
  highway = 40,    # fuel consumption in highway
  combined = 33)   # fuel consumption combined (city-and-hwy)
  # assing class
class(mazda3) <- "car"

mazda3
```

```{r}
  # add function to print() family
print.car <- function(x) {
  cat("Car\n")
  cat(sprintf('name: %s\n', x$name))
  cat(sprintf('make: %s\n', x$make))
  cat(sprintf('year: %s\n', x$year))
  invisible(x)
}

mazda3
```

```{r}
  # calculate travel miles
miles <- 
  function(car, fuel = 1, mpg = 'city') {
    stopifnot(class(car) == 'car')
    switch(mpg,
           'city' = car$city * fuel,
           'highway' = car$highway * fuel,
           'combined' = car$combined * fuel,
           car$city * fuel)
  }

miles(mazda3, fuel = 4, 'city')
miles(mazda3, fuel = 4, 'highway')
miles(mazda3, fuel = 4, 'combined')
```

```{r}
  # format travel miles
get_distance <- 
  function(car, fuel = 1, mpg = 'city') {
    distance <- miles(car, fuel = fuel, mpg = mpg) 
    cat(sprintf('A %s can travel %s miles\n',
                car$name, distance))
    cat(sprintf('with %s gallons of gas\n', fuel))
    cat(sprintf('using %s consumption', mpg))
  }

get_distance(mazda3, 4, 'city')
```

<br>

## 4.1.6 Example: Coffee Prices
```{r}
prices <- c(
  'americano' = 2, 
  'latte' = 2.75, 
  'mocha' = 3.45, 
  'capuccino' = 3.25)

class(prices); mode(prices)


print(prices)
```

```{r}
for (p in seq_along(prices)) {
  print(prices[p])
}

for (p in seq_along(prices)) {
  print(paste0(names(prices)[p], ' has price of $', prices[p]))
}

for (p in seq_along(prices)) {
  print(sprintf('%s has price of $%s', names(prices)[p], prices[p]))
}
```

```{r}
  # print without quote
for (p in seq_along(prices)) {
  noquote(sprintf('%s has price of $%s', names(prices)[p], prices[p]))
}

for (p in seq_along(prices)) {
  print(sprintf('%s has price of $%s', names(prices)[p], prices[p]), quote = FALSE)
}

for (p in seq_along(prices)) {
  cat(sprintf('%s has price of $%s\n', names(prices)[p], prices[p]))
}
```

